---
title: "Lab 06 - World TB"
author: "In Session"  
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: show
    theme: simplex
    highlight: tango
editor_options: 
  chunk_output_type: console
---

This report (add a brief description and the purpose of the report)
Demonstration of the __COMBINED__ template

<!--  Set the working directory to the repository's base directory; this assumes the report is nested inside of two directories.-->
```{r, echo=F, message=F, results = 'hide'}
cat("Working directory: ", getwd()) # turn on to test the location
# if the line above generates the project root, re-map by selecting 
# Session --> Set Working Directory --> To Source File location
```


```{r set_options, echo=F}
# cat("Working directory: ", getwd()) # turn on to test the location
report_render_start_time <- Sys.time()
# set options shared by all chunks
knitr::opts_chunk$set(
  results      = 'show',
  message      = FALSE,
  warning      = FALSE,
  comment      = NA,
  tidy         = FALSE,
  # dpi        = 400, # dots per inch,
  # out.width  = "650px", # pixels, this affects only the markdown, not the underlying png file.  The height will be scaled appropriately.
  fig.width    = 6, # inches
  fig.height   = 4, # inches
  fig.path     = 'figure-png-com/' # where figures are stored
)
echo_chunks    <- TRUE #Toggle for debugging.
message_chunks <- FALSE #Toggle for debugging.
options(width=100) # number of characters to display in the output (dflt = 80)
ggplot2::theme_set(ggplot2::theme_bw()) # common theme for all graphs
```

# Environment
<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-packages, message=message_chunks}
library(magrittr)  # pipes
library(dplyr)     # data wrangling
library(ggplot2)   # graphs
library(janitor)   # tidy data
library(tidyr)     # data wrangling
library(forcats)   # factors
library(stringr)   # strings
library(lubridate) # dates
library(wpp2019)

```

<!-- Load the sources.  Suppress the output when loading sources. --> 
```{r load-sources}
source("../../scripts/common-functions.R")
```

<!-- Load any Global functions and variables declared in the R file.  Suppress the output. --> 
```{r declare-globals}
# custom function for HTML tables
neat <- function(x, output_format = "html"){ 
  # knitr.table.format = output_format
  if(output_format == "pandoc"){
    x_t <- knitr::kable(x, format = "pandoc")
  }else{
    x_t <- x %>%
      # x %>%
      # neat() %>%
      knitr::kable(format=output_format) %>%
      kableExtra::kable_styling(
        bootstrap_options = c("striped", "hover", "condensed","responsive"),
        # bootstrap_options = c( "condensed"),
        full_width = F,
        position = "left"
      )
  } 
  return(x_t)
}
# Note: when printing to Word or PDF use `neat(output_format =  "pandoc")`
```



# Data
```{r load-data, message=FALSE, results='show'}
who <- who

library(wpp2019)
data("pop") #to load object from wpp2019

ds_geo <- read.csv("https://raw.githubusercontent.com/andkov/covid19-country-response/master/data-public/metadata/world-geography.csv")
```

## Tuberculosis

```{r tweak-data-1,echo = echo_chunks, results='show', message=message_chunks}
who1 <- who %>%
  pivot_longer(
      cols = new_sp_m014:newrel_f65
    , names_to       = "key"
    , values_to      = "case"
    , values_drop_na = TRUE
  )

who1
```

```{r tweak-data-2, echo = echo_chunks, results='show', message=message_chunks}
who2 <- who1 %>%
  mutate(
    key = stringr::str_replace(key, "newrel", "new_rel")
  )

who2
```

```{r tweak-data-3, echo = echo_chunks, results='show', message=message_chunks}
who3 <- who2 %>%
  separate(
    key, c("new", "type", "sexage"), sep = "_"
  )

who3
```

```{r tweak-data-4, echo = echo_chunks, results='show', message=message_chunks}
who4 <- who3 %>%
  select(-new) # Because all of the values are "new"
  # select(-c("iso2","iso3"))

who4
```

```{r tweak-data-5, echo = echo_chunks, results='show', message=message_chunks}
who5 <- who4 %>%
  separate(
    sexage, c("sex", "age"), sep = 1
  )

who5
ds_tb <- who5
```

## Population


```{r tweak-data-6, echo = echo_chunks, results='show', message=message_chunks}
var_select <- c("country_code", "name", seq(from = 2000, to = 2020, by = 5))

ds_pop <- pop %>% 
  select(
      all_of(var_select)
  ) %>% 
  tibble::as_tibble()
ds_pop
```
```{r}
a <- ds_tb %>% 
  distinct(country) %>% 
  mutate(
    country_tb = country
  )
b <- ds_pop %>% 
  distinct(name) %>% 
  rename(country = name) %>% 
  mutate(
    country_pop = country
  )
a
b

ab <- dplyr::full_join(a,b, by = "country")
# ab <- dplyr::bind_rows() 
ab


var_sel <- c("continent_code", "continent_name"
    ,"country_name", "country_label", "country_number"
    ,"iso2", "iso3"
    ,"year", "type", "sex", "age", "case"
    ,seq(from = 2000, to = 2020, by= 5))
ds_tb_geo_pop <- left_join(ds_tb, ds_geo,c("iso2" = "country_code2", "iso3" = "country_code")) %>% 
  left_join(ds_pop, c("country_number" = "country_code")) %>% 
  select(
    all_of(var_sel)
  )


ds_tb_geo_pop %>% glimpse()

# to do: 1. import pop data for males and females. 2. replace totals with gender breakdown and link to the final
# how would you approach this?
# look-up cleveland plots
```

Session Information {#session-info}
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
}
```
</details>

```{r session-duration, echo=FALSE}
report_render_duration_in_seconds <- round(as.numeric(difftime(Sys.time(), report_render_start_time, units="secs")))
```

Report rendered by `r Sys.info()["user"]` at `r strftime(Sys.time(), "%Y-%m-%d, %H:%M %z")` in `r report_render_duration_in_seconds` seconds.
